# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
#read data
```{r,cache=TRUE}
move.df<-read.csv(unz("activity.zip", "activity.csv"),header=TRUE, stringsAsFactors=FALSE)
move.df$date<-as.Date(move.df$date,"%Y-%m-%d")
```
Below is the first few rows of the read in data frame. 
```{r}
head(move.df)
```

## What is mean total number of steps taken per day?
After computing the average of the total steps per day, we plot the histogram.

```{r,cache=TRUE}
require(plyr)
perday.df<-ddply(move.df,.(date),summarise,sum=sum(steps,na.rm=T))
```

Below is a histgram of the total steps per day.
```{r histogram,cache=TRUE}
hist(perday.df$sum)
```
Below is the average and median.

```{r}
mean(perday.df$sum)
median(perday.df$sum)
```

## What is the average daily activity pattern?

Here we compute the average steps per interval.
```{r}
require(plyr)

stepPerInterval.df<-ddply(move.df,.(interval),
                          summarise,avgstep=mean(steps,na.r=T))
```

Below is a histogram of the average steps per interval. 
```{r}
plot(stepPerInterval.df,typ="l")
````

The interval with the maximum number of steps is computed below. 
```{r}
max.interval<-stepPerInterval.df[which.max(stepPerInterval.df$avgstep),"interval"]

```
The interval is `r max.interval`.




## Imputing missing values
The total number of NA row are computed below. 
```{r}
NA.rows<-nrow(move.df[!complete.cases(move.df),])
```
The number of rows are `r NA.rows` .

We now will now fill in the missing values with the average of the interval in the entire dataset. A new data frame has been created.

```{r,message=FALSE}
require(Hmisc); require(plyr)
```

```{r}
inpune.df<-ddply(move.df,"interval",mutate,imputed.value=impute(steps,mean))

inpune.df<-inpune.df[order(inpune.df$date),]
inpune.df$imputed.value<-round(inpune.df$imputed.value,0)

move.up.df<-inpune.df[,c(4,2,3)]

names(move.up.df)[1]<-"step"

head(move.up.df)
```

Below is a histogram using the new data set with the missing values filled in. 

```{r}

perday.up.df<-ddply(move.up.df,.(date),summarise,sum=sum(step,na.rm=T))

hist(perday.up.df$sum)
```

We follow with by computing a new mean and median below.
```{r}
mean(perday.up.df$sum)
median(perday.up.df$sum)
```
    
By filling in the missing values, data set characterstics have changed markedly.  First, both the mean and the median have moved higher, with the mean being affected more drastically upwards.  Additionally, the historgram looks more normally distributed after we filled in the missing values in the data set. 

## Are there differences in activity patterns between weekdays and weekends?
Below we have created a new variable ("week") indicating whether the date is a weekend or weekday.

```{r,message=FALSE}
require(chron)
```
```{r}
move.up.df$week<-is.weekend(move.up.df$date)
move.up.df[which(move.up.df$week==TRUE),"week"]<-"weekend"
move.up.df[which(move.up.df$week==FALSE),"week"]<-"weekday"
head(move.up.df,4)
```

Below we compute and plot time series of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r,message=FALSE}
require(ggplot2)
```
```{r}
stepPerIntervalWeek.df<-ddply(move.up.df,.(interval,week),summarise,avg.steps=mean(step,na.rm=T))


qplot(interval,avg.steps,facets = week~. ,
      data=stepPerIntervalWeek.df,geom="line")

```


